---
title: "readr$data.table"
author: "Manojit Sarkar"
date: "25 July 2018"
output: html_document
---

# Introduction
Till last chapter we have been using R base utils package which come preloaded with R .
Here we will be using two libraries readr and data.table

#readr

Written by Hadely Wickham fast and easy to use and very consistent.
so lets dive right in and install the package and loading it 

```{r}
#install.packages("readr")
library(readr)
```

Before we used read.csv()  now we will use read_csv() to read csv files -- read_csv returs a tibble (on tibble later-for now tibble is a supercharged version of a data frame; the prinout shows the classes of the coloumns.)

Similarly, instead of read.delim() we use read_tsv() (tsv--tab seperated values)


Let see some examples:

We will be using potatoes.csv 

```{r}
potatoes<-read_csv("C:/Users/gordon/Desktop/Data Camp/Datasets/potatoes.csv")
dim(potatoes)
```

Lets now see first 10 rows of potatoes

```{r}
head(potatoes,10)
```

Next lets read a .txt file seperated by tabs.
we will use read_tsv().
It doesn't have any header so we can add one to it.

```{r}
coloum_names <- c("area1", "temp1", "size1", "storage1", "method1","texture1", "flavor1", "moistness1")
potatoes1<-read_tsv("C:/Users/gordon/Desktop/Data Camp/Datasets/potatoes.txt",col_names = coloum_names)
head(potatoes1,10)
```

Now we will be using read_delim() and use delim argument which is similar to sep argument in read.table() 

we will use our potatoes.txt which is seperated by "\t" even we can use "" as the seperator


```{r}
coloum_names <- c("area1", "temp1", "size1", "storage1", "method1","texture1", "flavor1", "moistness1")
potatoes2<-read_delim("C:/Users/gordon/Desktop/Data Camp/Datasets/potatoes.txt",delim="\t",col_names = coloum_names)
head(potatoes2,10)
```

cool!!! lets now explore two more functions 

Through skip and n_max you can control which part of your flat file you're actually importing into R.

skip specifies the number of lines you're ignoring in the flat file before actually starting to import data.
n_max specifies the number of lines you're actually importing.

Say for example you have a CSV file with 20 lines, and set skip = 2 and n_max = 3, you're only reading in lines 3, 4 and 5 of the file.

Watch out: Once you skip some lines, you also skip the first line that can contain column names!

lets see this in action with an example

```{r}
coloum_names <- c("area1", "temp1", "size1", "storage1", "method1","texture1", "flavor1", "moistness1")
potatoes_fragment <- read_tsv("C:/Users/gordon/Desktop/Data Camp/Datasets/potatoes.txt", skip = 6, n_max = 5, col_names = coloum_names)
potatoes_fragment
```

As we can see above we hae skipped 6 rows nd then printed 5 rows..


We will now see how to use the col_type argument to change/update the class of coloumns

lets do a crazy thing let us change all the classes to character type .

```{r}
coloum_names <- c("area1", "temp1", "size1", "storage1", "method1","texture1", "flavor1", "moistness1")
potatoes_char <- read_tsv("C:/Users/gordon/Desktop/Data Camp/Datasets/potatoes.txt", col_types = "cccccccc",col_names = coloum_names)
str(potatoes_char)
```

 c-character, d-double, i-integer and l-logical _ skips the column as a whole.
 
 
 Now lets see other ways of specifying col_types---it can be done using collectors...collectors functions can be passed in a list() 

For this example we will use two collector functions:

col_integer(): the column should be interpreted as an integer.
col_factor(levels, ordered = FALSE): the column should be interpreted as a factor with levels.


We will be using our hd_nh.txt for this example : first we will import it normally do a summary on it then will use the col_class function using list()
so lets start

```{r}
hotdogs<-read_tsv("C:/Users/gordon/Desktop/Data Camp/Datasets/hd_nh.txt")
summary(hotdogs)
str(hotdogs)
```

now lets use col_types with list(collectors).

```{r}
fac <- col_factor(levels = c("Beef", "Meat", "Poultry"))
int <- col_integer()
hotdogs_factor<-read_tsv("C:/Users/gordon/Desktop/Data Camp/Datasets/hd_nh.txt",col_types = list(fac,int,int))
summary(hotdogs_factor)
str(hotdogs_factor)
```

Cool!!
