---
title: "importing from flat files--Utils"
author: "Manojit Sarkar"
date: "21 July 2018"
output: html_document
---
##INTRO

Lots of data comes in the form of flat files: simple tabular text files. Learn how to import all common formats of flat file data with base R functions.

A flat file database stores data in plain text format. Example csv -where csv stands for comma seperated values

## 1.CSV Files
state,capital,pop_mill,area_sqm
South Dakota,Pierre,0.853,77116
NewYork,Albany,19,746,54555
Oregon,Salem,3.970,98381

above is the example of a flat file in csv format.


But how to load this file into R -- we can do this using R base utils function read.csv()

lets load a csv file (you can download the csv file from my datasets folder)

```{r}
hr<-read.csv("C:/Users/gordon/Desktop/Data Camp/datasets/HR Analytics.csv")
head(hr,10)

```

in the path put your path for the csv file.

```{r}
str(hr)
```




Now here we can observe that all our string being converted into factor, if we dont want our strings to converted into factor we use the folloing code.


```{r}
hr1<-read.csv("C:/Users/gordon/Desktop/Data Camp/datasets/HR Analytics.csv",stringsAsFactors = FALSE)
head(hr1,10)
```
By default stringsAsFacor is set to TRUE.

```{r}
str(hr1)
```

As we can see the strings are not converted to factor here.

##2 Tab-delimited file

Are files seperated by "tabs"

Exapmle:
state   capital   pop_mill    area_sqm
South Dakota    Pierre    0.853   77116
New York    Albany    19.746    54555
Oregon    Salem   3.970   98381



```{r}
hotdog<-read.delim("C:/Users/gordon/Desktop/Data Camp/datasets/hot_dog.txt")
hotdog
```

Lets summarize the table

```{r}
summary(hotdog)
```

Great!!

## Exoctic Seperators 

data seperated by /, \ , # , * ,""etc etc

example:

state/capital/pop_mill/area_sqm
South Dakota/Pierre/0.853/77116
New York/Albany/19.746/54555
Oregon/Salem/3.970/ 98381

lets read the same hot_dog,text file with read.table

```{r}
hotdog1<-read.table("C:/Users/gordon/Desktop/Data Camp/datasets/hot_dog.txt",sep="",header=T)
hotdog1
```

Note: header in read.table is FALSE by Default.


now suppose there is a file without header let see how to add a header then.

```{r}
hotdog2<-read.table("C:/Users/gordon/Desktop/Data Camp/datasets/hd_nh.txt",sep="\t",header=F,col.names = c("type","calories","sodium"))
hotdog2
```


Select the hot dog with the least calories:

```{r}
leastcalories<-hotdog2[which.min(hotdog2$calories),]
leastcalories
```


Select the hot dog with the maximum sodium:

```{r}
maxsodium<-hotdog2[which.max(hotdog2$sodium),]
maxsodium
```


We can assign the class to the coloumns using colClasses

```{r}
hotdog3<-read.delim("C:/Users/gordon/Desktop/Data Camp/datasets/hd_nh.txt",header=F,col.names = c("type","calories","sodium"),colClasses = c("factor","NULL","numeric"))
str(hotdog3)
View(hotdog3)
```

Important point before we close this 

read.table() is the main function 
and 
read.csv ,read.delim are wrapper functions of read.table.




Thts all for this chapter!!! Meet u in the other one !!! Till then "May the force be with you!!!""