---
title: "tidyverse_datvisualisation"
author: "Manojit Sarkar"
date: "17 July 2018"
output: html_document
---

#Lets start!

## Data Visualisation

we start of with loading the ggplot 2 library / or the tidyverse library

Here i will be just loading the ggplot2 library 
note :- if the libraries arre missing you can download it by typing the following 

install.packages("packagename")


```{r}
library(ggplot2)
```

we will be using the same dataset as before the gapminder data set

if not installed : install it with the following line of code

```{r}
#install.packages("gapminder")
```

lets load the package

```{r}
library(gapminder)
```


Then install the dplyr package



```{r}
library(dplyr)
```
We will be working with the data pertaining to the year 2007 of gapminder and store it in a new data set gapminder_2007

For this we will be using the dplyr package :: filter verb

```{r}
gapminder_2007<-gapminder%>%
  filter(year==2007)
```

let us inspect the new dataset
```{r}
gapminder_2007
```

Cool!! We are now ready with our data set to carry out visualisation.


Lets do our fist visualisation usiing the ggplot2::ggplot

```{r}
ggplot(gapminder_2007,aes(x=gdpPercap,y=lifeExp))+geom_point()
```

Wow!!! isn't that a wonderful scatter plot.

We just plotted a scatter plot between gdpPercap in x axis and lifeExp in the Y  axis.










