---
title: "Tidyverse_data Camp"
author: "Manojit Sarkar"
date: "16 July 2018"
output: html_document
---

```{r}
#install.packages("gapminder"")
```
```{r}
library(gapminder)

```

Use str() on gapminder to know about class details and observation size of gapminder.

```{r}
str(gapminder)
```
 it shows there are 1704 rows with 6 coloumns
 with vabriables 
 
 country <- a factor data type with 155 levels<-categorical variable
 continent <- a factor with 5 levels<-categorical Variable
 year <- int (discrete variable)
 lifeExp<-numeric<-continous variable
 pop<- int (discrete)
 gdpPerCap<-num <-continous variable
 
```{r}
gapminder
```
 Just to see in the first 10 data
 
 
```{r}
#install.packages("dplyr")
```
```{r}
library(dplyr)
```
 
this is the tool for transforming the data such as filter,sorting and summarizing it.

## dplyr Verbs

Everytime we use a verb on a data set we have to use a pipe "%>%"
It says take whatever is before it and feed it to the next step

<B>1>filter() </B>

```{r}
gapminder%>%
  filter(year==2007)
```

Lets now find out the 2007 records for India

```{r}
gapminder%>%
  filter(year==2007,country=="India")
```

WoW ! It's so easy.

<B>2>arrange()</B>

Its shorts a table based on a variable

```{r}
gapminder%>%
  arrange(gdpPercap)
```

It arranges the data set in ascending order of gdpPercap. the data shows Democrartic Republic of congo had the lowesest gdp in the year 2002.

```{r}
gapminder%>%
  arrange(desc(gdpPercap))
```

with the "desc" function we change the default of arrange verb to descending order.

But looking at all countries for all the years might not be useful so lets look at countries arranged in order of gdpPercap for the year 2007

```{r}
gapminder%>%
  filter(year==2007)%>%
  arrange(gdpPercap)
```

Cool isn't it.

The order of verbs is not important as long as it doesn't break the logic.

what about another example : arrange in descending order of population for the year 1957

```{r}
gapminder%>%
  arrange(desc(pop))%>%
  filter(year==1957)
```

<B>3>mutate()</B>

It changes the variables in your dataset ,based on other ones or add a new variable also based on the existing variables.

First we will change an existing variable population. we will make population divide by a million.


```{r}
gapminder%>%
  mutate(pop=pop/1000000)
```

As we can see, the pop values are now in millions.


Now lets add a new variable to the dataset

Lets says we make a gdp variable which is gdppercapita multiplied by population

```{r}
gapminder%>%
  mutate(gdp=gdpPercap*pop)
```


Jinx! A new cloumn with gdp data. 

Now a big task!! Lets put together all the verbs we learned and find out the gdp of countries in the year 2007 in the descending order of gdp


```{r}
gapminder%>%
  mutate(gdp=gdpPercap*pop)%>%
  filter(year==2007)%>%
  arrange(desc(gdp))
```

Great!!

One more example

lets have a new variable lifeExpMonths which is nothing but life expectancy in months for the year 2007 and arrange it in descendending order of the new variable 

```{r}
gapminder%>%
  filter(year==2007)%>%
    mutate(lifeExpMonths=lifeExp*12)%>%
      arrange(desc(lifeExpMonths))
```

Magic!!!



